---
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    pandoc_args: ['-F', 'zotref', '-F', 'pandoc-citeproc']
title: Antidepressant Response and Side Effects in the GLAD Data
author: Yuhao Lin
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NA,
  prompt = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r clear global environment, include = FALSE}
remove(list = ls())
source("./analysis.r")
```

```{r rsetup, include=FALSE}
library(knitr)
opts_chunk$set(
  results = "asis",
  comment = NA,
  prompt = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

library(summarytools)
st_options(
  plain.ascii = FALSE,
  style = "rmarkdown",
  footnote = NA,
  subtitle.emphasis = FALSE
)
st_css()
```

## **Overview of the Measures Used in the Current Analyses**

# Medication Questionnaire

Measures specific to a medication:

**Medication History**: "Have you ever taken..."

  - With focuses on antidepressants and mood stablizers

**Effectiveness**: "How well does/did each antidepressant work for you?"

  - One three-point Likert scale item from 0 to 2

**Side Effects**


# Medication Questionnaire

Measures not specific to a medication:

**Best Aspects**: "What were the best aspects of taking the antidepressant(s)?"

  - Allows for multiple options, including items such as "Improved relationships", "Relief of depressive symptoms"

**Benefit**: "Overall, how would you rate the benefits of taking antidepressants?"

  - One five-point Likert scale item from 1 to 5

# Other Questionanires

**Patient Health Questionnaire (PHQ)-9**

**General Anxiety Disorder (GAD)-7**

**Standardized Assessment of Severity of Personality Disorder (SASPD)**

All measure current symptoms

**BMI**

# To be included

**Adapted CIDI Psychosis Module**

  - Having issues computing scores for this one

**Family History**

**Smoking status and physical conditions?**

# Computed variables used in the analyses

**Mean side effects**

  - Average number of side effects experienced across different
    medications

  - Discard some of the most common ones and rerun analyses?

**Mean effectiveness**

  - Average rating of effectiveness across different medications.

**Number of antidepressants taken in each class**

  - The classification is based on binding profile to receptors.

**Number of best aspects**

  - The number of items participant selected in the best aspect question.

# Descriptives

### All
```{r all, results = "asis"}
all
```

### Female
```{r female, results = "asis"}
female
```

### Male
```{r male, results = "asis"}
male
```

```{r cor, fig.width = 10, fig.height = 10, fig.align = "center"}
ggcorrplot(cor_mat,
  hc.order = TRUE, type = "lower",
  lab = TRUE,
  title = "Raw Correlations Between Items"
)
```

# OLS1

Mean side effects as response, benefits explanatory.
```{r}
tab_model(sef_ben_1,
  pred.labels = c(
    "Intercept", labels[
      !labels %in% c(
        "Mean Side Effects",
        "Mean Effectiveness",
        "Number of Best Aspects"
      )
    ]
  ),
  dv.labels = "Side Effects"
)
```

The residuals don't look right, which is related to the distribution of the
response.
```{r}
diag_ols_1[[1]]
diag_ols_1[[2]]
diag_ols_1[[3]]
diag_ols_1[[4]]
qplot(dat$mean_sef,
  geom = "histogram", binwidth = 1,
  xlab = "Mean Side Effects"
)
```

Try a simple transformation.
```{r}
tab_model(sef_ben_2,
  pred.labels = c(
    "Intercept", labels[
      !labels %in% c(
        "Mean Side Effects",
        "Mean Effectiveness",
        "Number of Best Aspects"
      )
    ]
  ),
  dv.labels = "log(Side Effects + 1)"
)
```

Unsatisfactory.
```{r}
diag_ols_2[[2]]
diag_ols_2[[3]]
diag_ols_2[[4]]
```

We then try to look for a GLM as simple as possible to model the data.
Proportional odds model seems to be a good choice if we treat "Benefit" as
response and model it as an ordinal variable. Since "Benefit" only takes five levels.

**Benefit**: "Overall, how would you rate the benefits of taking antidepressants?"

  - One five-point Likert scale item from 1 to 5

The idea of a proportional odds model is to use a cummulative link function

Suppose that the reponse has $J$ levels, we have for the response of each
participant $Y_{i}$ and level $j = 2, 3, ..., J$ (The first level is used
as reference)

$P(Y_{i} \le j) = G(\theta_{j} + \sum \beta x)$

So we are to fit a different intercept \theta_{j} but the same \betas to each level.
To express the right hand side as a linear combination of $x$s, we can use

$logit(P(Y_{i} \le j)) = log(\frac{P(Y_{i} \le j)}{P(Y_{i} > j)}$

So we can write

$log(\frac{P(Y_{i} \le j)}{P(Y_{i} > j)}) = \theta_{j} + \sum \beta x$

So fitting this model amounts to fitting $J-1$ logistic regression models with different intercepts.

```{r}

attributes <- attributes(plor_fit$model)
tab_model(plor_fit,
  pred.labels = c(
    "Intercept (<= 2)",
    "Intercept (<= 3)",
    "Intercept (<= 4)",
    "Intercept (<= 5)",
    labels[attributes$names]
  ),
  dv.labels = "Benefit"
)
```

Discard the unsignificant predictors and refit.
```{r}
attributes <- attributes(plor_fit2$model)
tab_model(plor_fit2,
  pred.labels = c(
    "Intercept (<= 2)",
    "Intercept (<= 3)",
    "Intercept (<= 4)",
    "Intercept (<= 5)",
    labels[attributes$names]
  ),
  dv.labels = "Benefit"
)
```

Also need to test the proportional odds assumption that all \betas are equal.

```{r}
bant_names <- c(
  "Omnibus",
  labels[
    !labels %in% c(
      "Benefit",
      "Age",
      "Mean Effectiveness",
      "Benefits",
      "Number of Best Aspects",
      "Current Anxiety",
      "NASSA",
      "SNRI",
      "SERO",
      "IRR",
      "NORA",
      "ALPHA1",
      "NRI"
    )
  ]
)
dimnames(bplor2)[[1]] <- bant_names
dimnames(bplor2)[[2]][3] <- "P value"
kable(bplor2, digits = 2)
```

The test is significant for MAO. Need to look in to detail by plotting the
model.  Potentially refit with a partial proportional odds model where some
\betas are free to varied between levels.

Now we turn to the mean effectiveness variable.

**Effectiveness**: "How well does/did each antidepressant work for you?"

  - One three-point Likert scale item from 0 to 2

**Mean effectiveness**

  - Average rating of effectiveness across different medications.

```{r}
qplot(dat$mean_eff,
  geom = "histogram", binwidth = 0.1,
  xlab = "Mean Effectiveness"
)
```

The distribution has "jumps" persumably because the variable is computed
from an ordinal item. We grouped the values by dividing into four
intervals, breaks of which are at each 0.5 increase, then fit with a
proportional odds model.

```{r}
attributes <- attributes(plor_fit_eff1$model)
tab_model(plor_fit_eff1,
  pred.labels = c(
    "Intercept (<= 1)",
    "Intercept (<= 1.5)",
    "Intercept (<= 2)",
    labels[attributes$names]
  ),
  dv.labels = "Grouped Mean Effectiveness"
)
```

Now turn to the "number of benefits" variable.

**Best Aspects**: "What were the best aspects of taking the antidepressant(s)?"

  - Allows for multiple options, including items such as "Improved relationships", "Relief of depressive symptoms"

**Number of best aspects**

  - The number of items participant selected in the best aspect question.

```{r}
attributes <- attributes(sef_n_best$model)
tab_model(sef_n_best,
  pred.labels = c(
    "Intercept",
    labels[attributes$names]
  ),
  dv.labels = "Number of Best Aspects"
)
```

# To be completed

* Association to each of the best aspect question

* Association to having to stop because of side effects

* Association to each of the worst aspect question

* Association between two further symptom questions - expect negative
  association
