---
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    pandoc_args: ['-F', 'zotref', '-F', 'pandoc-citeproc']
title: Untitled
subtitle: Untitled
author: Yuhao Lin
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = NA,
  prompt = FALSE,
  cache = FALSE,
  include = FALSE
)
```

```{r clear global environment, include = FALSE}
remove(list = ls())
```

```{r loading packages, include = FALSE}
library(tidyverse)
library(Hmisc)
library(sjPlot)
sef_options <- readLines("./sef_options.txt")
sef_optnames <- readLines("./sef_optnames.txt")
med_names <- readLines("./med_names.txt")
anxi <- c("Diazepam", "Lorazepam", "Clonazepam", "Nitrazepam")
sti <- c("Atomexatine", "Guanfacine", "Lisdexamfetamine", "Methylphenidate", "Dexamfetomine")
ane <- c("Topiramate", "Zonisamide")
ano <- c("Orlistat")
and <- c("Metformin", "Sibutramine")
slp <- c("Temazepam")
```

```{r}
dat <- read_csv("./MED_Renamed.csv") %>%
  rename(Sex = MED.SEX.1.0)
dat["Sex"] <- factor(dat[["Sex"]],
  levels = c(1, 0),
  labels = c("Female", "Male")
)
palette <- c("#efc00b", "#b7dee8")

is_rowna <- function(data) {
  apply(data, 1, function(x) {
    all(is.na(x))
  })
}

get_medname <- function(varname) {
  short_mednames <- str_split(varname, pattern = "\\.") %>%
    map_chr(nth, 3) %>%
    tolower()
  name <- map(short_mednames, function(x) {
    # We only need the first match where the first six letters
    # (not six letters in the middle) are matched.
    name <- grep(x, med_names, v = T, ignore.case = T)[1]
    if (is.na(name)) {
      return(x)
    } else {
      return(name)
    }
  })
  return(name)
}
```

```{r}
diagnoses <- c(
  "MDD", "BIP", "GAD", "PAN", "OCD", "SPE",
  "BDD", "PREM", "MENO", "PAIN", "SMOK",
  "SLEEP", "RLS", "EJAC", "ADHD", "BN",
  "BED", "OTHER", "NOTSURE", "PTSD", "DONTKNOW", "SCZA", "SCZ", "PSY",
  "EPIL", "IRR", "TIC", "NARC", "AS"
)

diag_cols <- map(diagnoses, function(x) {
  unique(grep(paste0("DX.", x), colnames(dat), v = T))
}) %>% unlist()

valid_n <- sum(!is_rowna(dat[diag_cols]))
na <- sum(is_rowna(dat[diag_cols]))

female_n <- nrow(dat[!is_rowna(dat[diag_cols]) & dat[["Sex"]] == "Female", ])
male_n <- nrow(dat[!is_rowna(dat[diag_cols]) & dat[["Sex"]] == "Male", ])

diagnoses_names <- c(
  "Depression", "Bipolar disorder", "Generalized anxiety disorder",
  "Panic disorder", "Obsessive compulsive disorder", "Specific phobia",
  "Body dysmorphic disorder", "Premenstrual dysphoric mood disorder",
  "Menopausal symptoms", "Chronic pain", "Quitting smoking",
  "Sleep problems", "Restless legs syndrome", "Premature ejaculation",
  "ADD/ADHD", "Bulimia nervosa", "Binge-eating disorder", "Other",
  "I'm not sure", "PTSD", "Don't know", "Schizoaffective disorder",
  "Schizophrenia", "Psychosis", "epilepsy", "Irritability/agitation",
  "Tics", "Narcolepsy", "Appetite suppression"
)

n_diag_female <- map_dbl(diagnoses, function(x) {
  sum(as.numeric(rowSums(dat[dat[["Sex"]] == "Female", grep(x, colnames(dat))], na.rm = T) >= 1))
}) %>%
  cbind("Female") %>%
  cbind(diagnoses_names)

n_diag_male <- map_dbl(diagnoses, function(x) {
  sum(as.numeric(rowSums(dat[dat[["Sex"]] == "Male", grep(x, colnames(dat))], na.rm = T) >= 1))
}) %>%
  cbind("Male") %>%
  cbind(diagnoses_names)

n_diag <- rbind(n_diag_female, n_diag_male)

n_diag_df <- tibble(Count = as.numeric(n_diag[, 1]), Sex = factor(n_diag[, 2]), Diagnosis = n_diag[, 3]) %>%
  group_by(Sex) %>%
  arrange(desc(Count))
```

```{r diagnosis count, include = TRUE, fig.width = 9, fig.height = 9}
ggplot(
  data = n_diag_df,
  aes(x = reorder(Diagnosis, Count), y = Count, fill = Sex)
) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = palette) +
  scale_color_manual(values = palette) +
  theme(
    panel.grid.major.x = element_line(
      size = 0.5,
      linetype = "dashed",
      colour = "gray"
    ),
    axis.title.y = element_blank(),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_blank(),
    legend.position = "bottom"
  ) +
  labs(
    y = "Frequency",
    title = "Diagnoses Count",
    subtitle = paste("n = ", valid_n,
      "; NA = ", na,
      "; n(total) = ", nrow(dat),
      "; n(female) = ", female_n,
      "; n(male) = ", male_n,
      sep = ""
    ),
    fill = "Sex",
    size = 15
  ) +
  geom_text(
    aes(label = Count),
    position = position_dodge(width = 0.9),
    hjust = -0.25, color = "black", size = 3
  ) +
  coord_flip(clip = "off")
```

```{r side effect per category}
dat_sef <- dat[grep("\\.SEF.", colnames(dat))]

categories <- colnames(dat_sef) %>%
  str_split("\\.") %>%
  map_chr(nth, 2) %>%
  unique() %>%
  .[. != "NA"]

sef_all <- colnames(dat_sef) %>%
  str_split("\\.") %>%
  map_chr(nth, 6)

categories_all <- colnames(dat_sef) %>%
  str_split("\\.") %>%
  map_chr(nth, 2)

meds_all <- colnames(dat_sef) %>%
  str_split("\\.") %>%
  map_chr(nth, 3)

med_taken <- paste("MED", categories_all, meds_all, sep = ".") %>%
  .[. != "MED.NA.CLICK "] %>%
  unique()

per_sef <- function(category) {
  sef <- sef_all[grep(category, colnames(dat_sef))] %>%
    unique()

  sef_renamed <- sef_optnames[sef_all %in% sef] %>% unique()

  dat_med_male <- dat[c("Sex", med_taken)] %>%
    select(Sex, contains(category)) %>%
    filter(Sex == "Male") %>%
    select(-Sex) %>%
    select(-matches("NEVER|NONE"))

  total_ins_male <- sum(rowSums(dat_med_male, na.rm = T))

  sef_n_male <- map_dbl(sef, function(s) {
    sef_col <- paste0(category, ".*", s)
    dat_sef_male <-
      dat_sef[grep(sef_col, colnames(dat_sef))] %>%
      bind_cols(dat["Sex"]) %>%
      filter(Sex == "Male") %>%
      select(-Sex)
    sum(colSums(dat_sef_male, na.rm = T))
  })

  sef_per_male <- cbind(sef_n_male / total_ins_male, sef_renamed, "Male")

  dat_med_female <- dat[c("Sex", med_taken)] %>%
    select(Sex, contains(category)) %>%
    filter(Sex == "Female") %>%
    select(-Sex) %>%
    select(-matches("NEVER|NONE"))

  total_ins_female <- sum(rowSums(dat_med_female, na.rm = T))

  sef_n_female <- map_dbl(sef, function(s) {
    sef_col <- paste0(category, ".*", s)
    dat_sef_female <-
      dat_sef[grep(sef_col, colnames(dat_sef))] %>%
      bind_cols(dat["Sex"]) %>%
      filter(Sex == "Female") %>%
      select(-Sex)
    sum(colSums(dat_sef_female, na.rm = T))
  })

  sef_per_female <- cbind(sef_n_female / total_ins_female, sef_renamed, "Female")
  sef_per <- rbind(sef_per_male, sef_per_female)
  sef_per_df <- tibble(
    "Percentage" = as.numeric(sef_per[, 1]),
    "SEF" = sef_per[, 2],
    "Sex" = sef_per[, 3]
  )
  return(sef_per_df)
}

category_titles <- c(
  "Antidepressants", "Mood Stablizers",
  "Stimulants", "Anti-epileptic",
  "Anti-obesity", "Anti-diabetics",
  "Anxiolytic", "Sleep"
) %>% setNames(categories)
```

This is the number of instances of a side effect divided by the number of
instances of taking medications belonging to a category.

* `r category_titles[3]`: `r sti`
* `r category_titles[4]`: `r ane`
* `r category_titles[5]`: `r ano`
* `r category_titles[6]`: `r and`
* `r category_titles[7]`: `r anxi`
* `r category_titles[8]`: `r slp`

```{r side effect per medication category, include = TRUE, fig.width = 9, fig.height = 9}
for (category in categories) {
  sef_type <- unique(per_sef(category)[["SEF"]])
  cut_factor <- cut2(seq_along(sef_type), m = 10)

  category_col <- grep(category, med_taken, v = T)
  category_valid_n <- sum(!is_rowna(dat[category_col]))
  category_na <- sum(is_rowna(dat[category_col]))

  category_female_n <-
    nrow(dat[!is_rowna(dat[category_col]) & dat[["Sex"]] == "Female", ])
  category_male_n <-
    nrow(dat[!is_rowna(dat[category_col]) & dat[["Sex"]] == "Male", ])

  for (i in levels(cut_factor)) {
    is <- which(cut_factor == i)
    sef_plotdat <- filter(
      per_sef(category),
      SEF %in% sef_type[is]
    )
    sef_plot <- ggplot(
      data = sef_plotdat,
      aes(x = SEF, y = Percentage, fill = Sex)
    ) +
      geom_bar(stat = "identity", position = position_dodge()) +
      scale_fill_manual(values = palette) +
      scale_color_manual(values = palette) +
      theme(
        panel.grid.major.x = element_line(
          size = 0.5,
          linetype = "dashed",
          colour = "gray"
        ),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour = "black", size = 12),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom"
      ) +
      labs(
        title = paste(
          "Percentage Suffering From Side Effects in",
          category_titles[category]
        ),
        subtitle = paste("n = ", category_valid_n,
          "; n(female) = ", category_female_n,
          "; n(male) = ", category_male_n,
          sep = ""
        ),
        fill = "Sex"
      ) +
      geom_text(
        aes(label = paste0(round(Percentage, digits = 3) * 100, "%")),
        position = position_dodge(width = 0.9),
        hjust = -0.25, color = "black", size = 3
      ) +
      coord_flip(clip = "off")
    print(sef_plot)
  }
}
```

Average in the sense that a participant can take multiple medications
(within a category) and we're averaging the number of side effects by
the number of medications taken.

```{r average side effects per category, include = TRUE, fig.width = 9, fig.height = 9}
avg_sef <- function(category) {
  sef <- sef_all[grep(category, colnames(dat_sef))] %>%
    unique()

  dat_med_male <- dat[c("Sex", med_taken)] %>%
    select(Sex, contains(category)) %>%
    filter(Sex == "Male") %>%
    select(-Sex)

  n_dx_male <- rowSums(dat_med_male, na.rm = T)

  tsef_col <- unlist(map(
    paste0(category, ".*", sef),
    grep, colnames(dat_sef)
  ))

  dat_tsef_male <-
    dat_sef[tsef_col] %>%
    bind_cols(dat["Sex"]) %>%
    filter(Sex == "Male") %>%
    select(-Sex)

  n_sef_male <- rowSums(dat_tsef_male, na.rm = T)

  avg_sef_male <- (n_sef_male / n_dx_male) %>%
    subset(!is.nan(.)) %>%
    setNames(rep("Male", length(.))) %>%
    enframe(name = "Sex")

  dat_med_female <- dat[c("Sex", med_taken)] %>%
    select(Sex, contains(category)) %>%
    filter(Sex == "Female") %>%
    select(-Sex)

  n_dx_female <- rowSums(dat_med_female, na.rm = T)

  dat_tsef_female <-
    dat_sef[tsef_col] %>%
    bind_cols(dat["Sex"]) %>%
    filter(Sex == "Female") %>%
    select(-Sex)

  n_sef_female <- rowSums(dat_tsef_female, na.rm = T)

  avg_sef_female <- (n_sef_female / n_dx_female) %>%
    subset(!is.nan(.)) %>%
    setNames(rep("Female", length(.))) %>%
    enframe(
      name = "Sex",
    )

  avg_sef <- bind_rows(avg_sef_female, avg_sef_male)
  return(avg_sef)
}

for (category in categories) {
  category_col <- grep(category, med_taken, v = T)
  category_valid_n <- sum(!is_rowna(dat[category_col]))

  category_female_n <-
    nrow(dat[!is_rowna(dat[category_col]) & dat[["Sex"]] == "Female", ])
  category_male_n <-
    nrow(dat[!is_rowna(dat[category_col]) & dat[["Sex"]] == "Male", ])

  avg_sefplotdat <- avg_sef(category)

  stats_sex <- avg_sefplotdat %>%
    group_by(Sex) %>%
    dplyr::summarize(
      grp.mean = mean(value),
      grp.sd = sd(value)
    )

  avg_sefplot <- ggplot(
    data = avg_sefplotdat,
    aes(x = value)
  ) +
    geom_histogram(aes(color = Sex, fill = Sex, alpha = 0.5),
      binwidth = 1,
      position = "identity"
    ) +
    scale_fill_manual(values = palette) +
    scale_color_manual(values = palette) +
    theme(
      panel.grid.major.y = element_line(
        size = 0.5,
        linetype = "dashed",
        colour = "gray"
      ),
      axis.text.x = element_text(colour = "black", size = 12),
      axis.text.y = element_text(colour = "black", size = 12),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_blank(),
      panel.background = element_blank(),
      legend.position = "bottom"
    ) +
    guides(alpha = FALSE, color = FALSE) +
    geom_vline(data = stats_sex, aes(xintercept = grp.mean, colour = Sex), size = 0.5, linetype = "dashed") +
    geom_vline(data = stats_sex, aes(xintercept = grp.mean + grp.sd, colour = Sex), size = 0.5, linetype = "dotted") +
    geom_vline(data = stats_sex, aes(xintercept = grp.mean - grp.sd, colour = Sex), size = 0.5, linetype = "dotted") +
    labs(
      x = "Average Numer of Side Effects",
      y = "Count",
      title = paste(
        "Average Numer of Side Effects For",
        category_titles[category]
      ),
      subtitle = paste0(
        "Dashed line represents mean and dotted lines represent  1 standard deviation\n",
        "n(female) = ", category_female_n,
        "; n(male) = ", category_male_n
      ),
      fill = "Sex"
    )
  print(avg_sefplot)
}
```

```{r effectiveness rating female, include = TRUE, fig.width = 9, fig.height = 9, warning = FALSE}
for (category in categories[1:2]) {
  dat_cat_eff <- dat %>%
    filter(Sex == "Female") %>%
    select(contains(category)) %>%
    select(contains("EFF")) %>%
    map_df(factor, levels = 0:2)

  order_by_n <- order(
    map_int(dat_cat_eff, ~ sum(!is.na(.x)))
  )
  dat_cat_eff <- dat_cat_eff %>%
    select(order_by_n)
  plot <- plot_stackfrq(dat_cat_eff,
    title = paste(
      category_titles[category],
      "Effectiveness Rating"
    ),
    axis.labels = get_medname(colnames(dat_cat_eff))
  )
  print(plot)
}
```

```{r effectiveness rating male, include = TRUE, fig.width = 9, fig.height = 9, warning = FALSE}
for (category in categories[1:2]) {
  dat_cat_eff <- dat %>%
    filter(Sex == "Male") %>%
    select(contains(category)) %>%
    select(contains("EFF")) %>%
    map_df(factor, levels = 0:2)

  order_by_n <- order(
    map_int(dat_cat_eff, ~ sum(!is.na(.x)))
  )
  dat_cat_eff <- dat_cat_eff %>%
    select(order_by_n)
  plot <- plot_stackfrq(dat_cat_eff,
    title = paste(
      category_titles[category],
      "Effectiveness Rating"
    ),
    axis.labels = get_medname(colnames(dat_cat_eff))
  )
  print(plot)
}
```

```{r}
dat_dx <- dat[grep("\\.DX.", colnames(dat))] %>%
  select(-matches("NEVER|NONE|OTHERRE|OTHERADDRE|OTHERMSTRE|OTHER")) %>%
  select(-matches("START|DUR"))

dx_uniq <- colnames(dat_dx) %>%
  str_split("\\.") %>%
  map_chr(nth, 5) %>%
  unique() %>%
  .[!. %in% c("DONTKNOW", NA)]

# The idea is, for each unique dx, find who indicates he has this dx and
# what medications are used for this dx. Then for each participant with
# this dx, find for each medication related to this dx (from the dx slice
# colnames), the number of sef from the sef slice (by matching the
# medication names), obtain the sum of the number of sef across the
# medications and divide this by the number of medications taken. Finally,
# this average numbers of side effects for participants across the
# medications taken for this diagnosis are averaged across the number of
# participants with this diagnosis to obtain the average number of side
# effects experienced "resulted" from a diagnosis.

# I'm a bit short of time to write code for this...

dx <- "MDD"
map(dx_uniq, function(dx) {
  i <- dat_dx %>%
    select(contains(dx)) %>%
    is_rowna() %>%
    !.

  n_med_male <- dat[i, c("Sex", med_taken)] %>%
    filter(Sex == "Male") %>%
    select(-Sex) %>%
    select(-matches("NEVER|NONE")) %>%
    rowSums(na.rm = T)

  total_ins_male <- sum(rowSums(dat_med_male, na.rm = T))

  n_dx_sef <- dat_sef[i, ] %>%
    bind_cols(dat["Sex"]) %>%
    filter(Sex == "Male") %>%
    # What to remove?
    select(-matches("NEVER|NONE")) %>%
    rowSums()
})
```

