---
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    pandoc_args: ['-F', 'zotref', '-F', 'pandoc-citeproc']
title: Investigating the Relationship Between Self-report Side Effect and Antidepressant Effectiveness
subtitle: Preliminary Analyses
author: Yuhao Lin
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r clear global environment variables, include = FALSE}
rm(list = setdiff(ls(), lsf.str()))
```

```{r import datasets, include = FALSE}
source("../../cleaning_scripts/med/MED.r")
```

```{r rsetup, include=FALSE}
library(knitr)
opts_chunk$set(
  results = "asis",
  comment = NA,
  prompt = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

library(summarytools)
st_options(
  plain.ascii = FALSE,
  style = "rmarkdown",
  footnote = NA,
  subtitle.emphasis = FALSE
)

st_css()

library(tidyverse)
library(forcats)
library(ggplot2)
library(GGally)
library(lindia)
library(sjPlot)
library(sjmisc)
library(sjPlot)
library(car)
library(broom)
library(kableExtra)
options(knitr.kable.NA = "")
library(vcd)
```

```{r extract the required columns, include = F}
add_names <- med.names[grep("MED.ADD", colnames(dat_med[items][posi.med]))] %>%
  .[-length(.)] # Last item is "I have never..."

add_eff <- dat_med[items][posi.eff][eff.names %in% add_names] %>%
  map_df(function(x) {
    x[x == -88] <- NA
    return(as.numeric(x) - 1)
  })

mean_eff <- rowSums(add_eff, na.rm = T) / dat_med$MED.ADD.N %>%
  as.numeric() %>%
  enframe(name = NULL)
names(mean_eff) <- "mean_eff"

add_sef <- dat_med[items][posi.sef][sef.names %in% add_names] %>%
  map_df(as.numeric)
mean_sef <- rowSums(add_sef, na.rm = T) / dat_med$MED.ADD.N %>%
  as.numeric() %>%
  enframe(name = NULL)
names(mean_sef) <- "mean_sef"

benefit <- dat_med[items][posi.ben][1]
names(benefit) <- "benefit"

dat <- bind_cols(
  dat_med[1:3],
  mean_eff, mean_sef,
  benefit
) %>%
  rename(sex = MED.SEX.1.0) %>%
  mutate(sex = as.numeric(sex)) %>%
  mutate(sex = haven::labelled(sex, c(Female = 1, Male = 0))) %>%
  mutate(sex = haven::as_factor(sex, levels = "labels"))

label_names <- c(
  "Sex", "Age",
  "Effectiveness", "Side Effects", "Benefit"
) %>%
  setNames(colnames(dat)[-1])
```

## Descriptive Statistics

* The "Benefit" variable refers to the answer to
  "Overall, how would you rate the benefits of taking antidepressants?".
  The answers range from 1 to 5.

* The "Side Effect" variable is the total number of side effects
  experienced divided by the number of medications (i.e. average number of
  side effects across different medications). This is preliminary since
  Gerome suggests we should discuss with the clinician to work out
  a better measure.

* The "Effectiveness" variable is the sum of the responses on
  "How well does/did each antidepressant work for you?" (ranging
  from 0 to 2) divided by the number of medications (i.e. average
  rating of effectiveness across different medications).

### All

```{r descriptives all}
summarytools::descr(dat %>%
  plyr::rename(label_names),
transpose = T,
stats = c(
  "mean", "med", "sd", "max",
  "min", "skewness"
), headings = FALSE
)
```

### Female

```{r descriptives female}
summarytools::descr(dat %>%
  filter(sex == "Female") %>%
  plyr::rename(label_names),
transpose = T,
stats = c(
  "mean", "med", "sd", "max",
  "min", "skewness"
), headings = FALSE
)
```

### Male

```{r descriptives male}
summarytools::descr(dat %>%
  filter(sex == "Male") %>%
  plyr::rename(label_names),
transpose = T,
stats = c(
  "mean", "med", "sd", "max",
  "min", "skewness"
), headings = FALSE
)
```

```{r check outliers, include = F}
# Checking outliers
freq(dat$mean_eff)
freq(dat$mean_sef)
freq(dat$benefit)
freq(dat$age)
freq(dat$sex)
# Seems fine.
```

## Checking distributions {.tabset .tabset-fade}

### Age
```{r age distribution, echo = F}
qplot(dat$age,
  geom = "histogram", binwidth = 1,
  xlab = "Age"
)
```

### Side Effects
```{r mean_sef distribution, echo = F}
qplot(dat$mean_sef,
  geom = "histogram", binwidth = 1,
  xlab = "Mean Number of Side Effects"
)
```

### Benefit
```{r benefit distribution, echo = F}
qplot(dat$benefit,
  geom = "histogram", binwidth = 1,
  xlab = "Benefit Rating"
)
```

### Effectiveness
```{r effectiveness distribution, echo = F}
qplot(dat$mean_eff,
  geom = "histogram", binwidth = 1,
  xlab = "Mean Effectiveness"
)
```

## Examining correlations
```{r correlations, fig.width = 45, fig.height = 45}
ggpairs(
  data = dat[c("sex", "age", "mean_sef", "benefit", "mean_eff")],
  lower = list(continuous = "smooth", combo = "box"),
  diag = list(continuous = "bar"),
  upper = list(combo = "blank", continuous = wrap("cor", size = 15)),
  columnLabels = c(
    "Sex", "Age",
    "Side Effects",
    "Benefit", "Effectiveness"
  ),
  axisLabel = "show"
) + theme_grey(base_size = 50)
```

## Gender Comparison
t-tests to compare variables across genders.
```{r gender comparison, echo = F}

test <- bind_rows(
  tidy(t.test(mean_eff ~ sex, data = dat, var.equal = F)),
  tidy(t.test(mean_sef ~ sex, data = dat, var.equal = F)),
  tidy(t.test(benefit ~ sex, data = dat, var.equal = F))
) %>%
  mutate(Variable = c("Side Effect", "Effectiveness", "Benefit")) %>%
  select(
    Variable, parameter, estimate, estimate1, estimate2, p.value,
    conf.low, conf.high
  )

kable(test,
  digits = 2,
  col.names = c(
    "", "df", "Difference", "Male", "Female",
    "p value", "CI low", "CI high"
  )
) %>%
  column_spec(1, bold = T)
```

## OLS Models

### Fitting OLS1

```{r ols1, echo = F}
sef_ben_1 <- lm(data = dat, mean_sef ~ benefit + sex + age)
tab_model(sef_ben_1,
  pred.labels = c(
    "Intercept", "Benefit",
    "Sex", "Age"
  ),
  dv.labels = "Side Effects"
)
```
Not much variance explained here. Is this what Gerome expects?
If I did not misunderstand he seemed to think side effects can
be a proxy for effectiveness, but here it seems not.

### Diagnostics OLS1

```{r ols1 assumptions}
diag_ols_1 <- plot_model(sef_ben_1, type = "diag")
print(diag_ols_1[[1]])
print(diag_ols_1[[2]])
print(diag_ols_1[[3]])
print(diag_ols_1[[4]])
gg_cooksd(sef_ben_1)

cook_dist <- cooks.distance(sef_ben_1)
n_influ <- sum(cook_dist > (4 / (nrow(dat) - 4 - 1)))
```

* Positively skewed residuals overall, also conditionally
at each level of `benefit`. I presume it's due to
the outcome variable `mean_sef` being also
positively skewed?

* Seems to be many influential points.
Discard and refit? Rule of thumb
with regard to Cook's distance is
points with distance larger than
$\frac{4}{n-p-1}$, where $p$ is the
number of parameters.

### Examining Influential Points
Let's look at the first 20 out of `r n_influ`.

```{r examining influential points}
influ_dat <- dat[complete.cases(dat[-4]), ][order(cook_dist, decreasing = T), ] %>%
  plyr::rename(label_names)
kable(influ_dat %>% select(-ID) %>% head(20), digits = 2)
```

Nothing too special, it seems? So we will keep them.

For issues with non-normal residuals,
perhaps do a transformation (let the response
variable be $log(Side Effects + 1)$) ?
Not so sure since Gerome wants it to be simple and descriptive,
and the untransformed model is certainly easier to interpret.

### Fitting OLS2

```{r ols2}
dat2 <- dat
dat2$log_mean_sef <- log(dat2$mean_sef + 1)
sef_ben_2 <- lm(data = dat2, log_mean_sef ~ benefit + sex + age)
tab_model(sef_ben_2,
  pred.labels = c(
    "Intercept", "Benefit",
    "Sex", "Age"
  ),
  dv.labels = "log(Side Effects + 1)"
)
```

### Diagnostics OLS2

```{r ols2 diagnostics}
diag_ols_2 <- plot_model(sef_ben_2, type = "diag")
print(diag_ols_2[[1]])
print(diag_ols_2[[2]])
print(diag_ols_2[[3]])
print(diag_ols_2[[4]])
gg_cooksd(sef_ben_2)
```

A bit better.

### A quick look at effectiveness and side effects
```{r effectiveness and side effects}
sef_eff_1 <- lm(data = dat, mean_sef ~ mean_eff + sex + age)
tab_model(sef_eff_1,
  pred.labels = c(
    "Intercept", "Effectiveness",
    "Sex", "Age"
  ),
  dv.labels = "Side Effects (Model 2)"
)
```

The fit is worse.

## Relationship of Effectiveness and Side Effect in Three Most Taken Medications

```{r retrieve columns for the most taken meds}
med_count <- dat_med[items][posi.med] %>%
  select(contains("ADD")) %>%
  colSums(na.rm = T)

most_names <- med.names[order(med_count, decreasing = T)] %>% head(3)
eff_most <- dat_med[items][posi.eff][eff.names %in% most_names] %>%
  map_df(function(x) {
    return(factor(x, levels = 0:2))
  }) %>%
  rename(
    eff_1 = MED.EFF.1.0,
    eff_2 = MED.EFF.2.0,
    eff_3 = MED.EFF.3.0
  )

sef_most <- dat_med[items][posi.sef][sef.names %in% most_names]

sef_1 <- rowSums(sef_most[grep("1.0", colnames(sef_most))],
  na.rm = T
) %>%
  enframe(name = NULL) %>%
  setNames("sef_1")

sef_2 <- rowSums(sef_most[grep("2.0", colnames(sef_most))],
  na.rm = T
) %>%
  enframe(name = NULL) %>%
  setNames("sef_2")

sef_3 <- rowSums(sef_most[grep("3.0", colnames(sef_most))],
  na.rm = T
) %>%
  enframe(name = NULL) %>%
  setNames("sef_3")

dat_most <- bind_cols(
  dat_med[1:3],
  eff_most, sef_1,
  sef_2, sef_3
) %>%
  rename(sex = MED.SEX.1.0) %>%
  mutate(sex = as.numeric(sex)) %>%
  mutate(sex = haven::labelled(sex, c(Female = 1, Male = 0))) %>%
  mutate(sex = haven::as_factor(sex, levels = "labels"))
```

### Anova Models

We include the interaction terms of the continuous variable and
the categorical variables to test the assumption of homogeneity of
regression slopes.

#### **Citalopram**

```{r anova1}
anova1 <- aov(sef_1 ~ eff_1 + sex + age + age:sex + age:eff_1, data = dat_most)

anova1_tab <- Anova(anova1, type = 3)
anova1_tab$pes <- c(anova1_tab$"Sum Sq"[-nrow(anova1_tab)], NA) / (anova1_tab$"Sum Sq" + anova1_tab$"Sum Sq"[nrow(anova1_tab)])

colnames(anova1_tab) <- c("SS", "df", "$F$", "$p$", "partial $\\eta^2$")
rownames(anova1_tab) <- c(
  "(Intercept)",
  "Effectiveness",
  "Sex",
  "Age",
  "Sex x Age",
  "Effectiveness x Age",
  "Residuals"
)

levene_test1 <- leveneTest(sef_1 ~ sex * eff_1, data = dat_most)
colnames(levene_test1) <- c("df", "$F$", "$p$")
rownames(levene_test1) <- c("Group", "Observations")

kable(anova1_tab,
  digits = 2,
) %>% column_spec(1, bold = T)

kable(levene_test1,
  digits = 2,
  caption = "Levene's Test"
)
```

#### **Sertraline**

```{r anova2}
anova2 <- aov(sef_2 ~ eff_2 + sex + age + age:sex + age:eff_2, data = dat_most)

anova2_tab <- Anova(anova2, type = 3)
anova2_tab$pes <- c(anova2_tab$"Sum Sq"[-nrow(anova2_tab)], NA) / (anova2_tab$"Sum Sq" + anova2_tab$"Sum Sq"[nrow(anova2_tab)])

colnames(anova2_tab) <- c("SS", "df", "$F$", "$p$", "partial $\\eta^2$")
rownames(anova2_tab) <- c(
  "(Intercept)",
  "Effectiveness",
  "Sex",
  "Age",
  "Sex x Age",
  "Effectiveness x Age",
  "Residuals"
)


levene_test2 <- leveneTest(sef_2 ~ sex * eff_2, data = dat_most)
colnames(levene_test2) <- c("df", "$F$", "$p$")
rownames(levene_test2) <- c("Group", "Observations")

kable(anova2_tab,
  digits = 2,
) %>% column_spec(1, bold = T)

kable(levene_test2,
  digits = 2,
  caption = "Levene's Test"
)
```

#### **Fluoxetine**

```{r anova3}
anova3 <- aov(sef_3 ~ eff_3 + sex + age + age:sex + age:eff_3, data = dat_most)

anova3_tab <- Anova(anova3, type = 3)
anova3_tab$pes <- c(anova3_tab$"Sum Sq"[-nrow(anova3_tab)], NA) / (anova3_tab$"Sum Sq" + anova3_tab$"Sum Sq"[nrow(anova3_tab)])

colnames(anova3_tab) <- c("SS", "df", "$F$", "$p$", "partial $\\eta^2$")
rownames(anova3_tab) <- c(
  "(Intercept)",
  "Effectiveness",
  "Sex",
  "Age",
  "Sex x Age",
  "Effectiveness x Age",
  "Residuals"
)

levene_test3 <- leveneTest(sef_3 ~ sex * eff_3, data = dat_most)
colnames(levene_test3) <- c("df", "$F$", "$p$")
rownames(levene_test3) <- c("Group", "Observations")

kable(anova3_tab,
  digits = 2,
) %>% column_spec(1, bold = T)

kable(levene_test3,
  digits = 2,
  caption = "Levene's Test"
)
```

None of these are significant. Also Levene's tests indicate
inhomogeneity of variances for all variables
(higher probability of false positives).
